---
title: "igraph test"
author: "Lucy Andrews"
date: "9/20/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
# load packages
suppressPackageStartupMessages({
  library(tidyverse);
  library(here);
  library(sf);
  library(glue);
  library(stringr);
  library(igraph);
  library(networkD3);
  library(mapview);
  mapviewOptions(fgb = FALSE)
  })
```

To use this in a watershed, replace the `up_down_df` below with your two-column adjacency/edge dataframe.

Be sure that column headings match (`up`, `down`) or change the names of column headings throughout this script to match your input dataframe.

## Get Flowline Data

```{r getdata}
# set up two-column edge dataframe
flowlines <- read_rds(here("data_output/final_flowlines_w_full_nhd_vaa.rds"))

# reduce fields for plotting purposes
flowlines_map <- flowlines %>% select(id, comid, contains("seq"), hydroseq, gnis_name, areasqkm:divdasqkm, shape_length, streamorde, streamorder_map, streamcalc, geom)

# fix flowlines comid to factor in correct order
flowlines_map <- flowlines_map %>%
  mutate(comid_f = factor(as.character(comid),
                          levels=c(3917198, 3917200, 3917948,
                                   3917950, 3917244, 3917946)))

# drop sinks (isolated channels)
sinks <- c(3917228, 3917212, 3917214, 3917218, 3917220,
           3917960, 3917958, 3917276, 3917278, 3917274, 
           3917282, 3917284, 3917286, 3917280, 3917268,
           3917256, 3917250, 3917272, 3917956)

flowlines_map_trim <- flowlines_map %>% 
  filter(!comid %in% sinks)

mapview(flowlines_map_trim)

```

## Create adjacency matrix

Columns contain vertices and each full row represents an edge (from a vertex to vertex). 

For a watershed, this will be a *to-from* dataframe with hydroseq values associated with nodes. We can use NAs to flag terminal values for ease of comparison with network image, but can be filtered out to create the igraph. This drops vertices that do not compose edges.

### Example Network

```{r getadj-ex}
# demo
up_down_df <- tribble(
  ~up, ~down,
  "Q", NA,
  "P", "N",
  "O", "N",
  "N", "M",
  "M", NA,
  "L", "G",
  "K", "G",
  "J", "G",
  "I", "E",
  "H", "E",
  "G", "C",
  "F", "B",
  "E", "B",
  "D", "A",
  "C", "A",
  "B", "A",
  "A", NA,
  "R", "Q",
  NA, "P",
  NA, "O",
  NA, "L",
  NA, "K",
  NA, "J",
  NA, "I",
  NA, "H",
  NA, "F",
  NA, "D",
  NA, "R",
  "S", "T",
  "T", NA,
  "R", "S",
  "U", NA,
  NA, "U"
  ) %>%
  filter(!is.na(up) & !is.na(down))

# tst
(p <- simpleNetwork(up_down_df, height = "400px", 
                    width = "400px", 
                    fontSize = 16, 
                    fontFamily = "serif",
                    nodeColour = "darkblue", 
                    linkColour = "steelblue",
                    opacity = 0.9, zoom = TRUE, charge = -40))
```

### Little Shasta Network

```{r getadj-lsh}

library(nhdplusTools)

# get nodes: Start
starts <- get_node(flowlines_map_trim, "start") %>%
  mutate(hydroseq=flowlines_map_trim$hydroseq) %>% 
  arrange(hydroseq) %>% 
  mutate(us_ord = rev(seq(1:nrow(.))))

# get midpoints
midpts <- st_point_on_surface(flowlines_map_trim)

mapview(starts, zcol="hydroseq", legend=FALSE,
        layer.name="Starts") + 
  mapview(flowlines_map_trim, legend=FALSE) +
  mapview(flowlines_map) +
  mapview(midpts, col.regions="green", cex=3)

# get matrix
up_down_lsh <-
  midpts %>% st_drop_geometry() %>% 
  select(up = hydroseq,
         down = dnhydroseq) %>%
  # fix the downstream end: 10020426 to NA
  mutate(down = case_when(
    up == 10020426 ~ 0, # this is OUTLET
    TRUE ~ down)) %>% 
  filter(!is.na(down)) # make sure no NAs

# spur at ds end is: 10135201, penultimate seg: 10020862

# tst
(p <- simpleNetwork(up_down_lsh, Source = "up", Target = "down", height = "400px", 
                    width = "400px", 
                    fontSize = 16, 
                    fontFamily = "serif",
                    nodeColour = "darkblue", 
                    linkColour = "steelblue",
                    opacity = 0.9, zoom = TRUE, charge = -40))

```

## Identify Termini

Need to update and identify the termini to create the "edges" of the network.

### Upstream Ends

We want to identify the upstream ends, which is where paths stop. To run the other direction (upstream --> downstream), we can change what is considered "upstream" or reverse the `all_simple_paths` call below.

This way of doing it uses an edge data frame, but we could also specify upstream ends if needed, and they don't have to be termini (e.g. you can stop mid-network).

```{r usends}

# this identifies the upstream "ends" (termini)
upstream_ends <- up_down_lsh %>%
  filter(!up %in% down) %>%
  pull(up) %>%
  unique()
```

### Downstream Ends

Next we can identify downstream ends. The upstream ends are the places paths stop, so to run the other direction (upstream --> downstream), we can change  what is considered "downstream" or reverse the `all_simple_paths` call below. 

```{r dsends}

downstream_ends <- up_down_lsh %>%
  filter(!down %in% up) %>%
  pull(down) %>%
  unique()

```

## Convert to Matrix

```{r tomatrix}

# turn edge dataframe into matrix for igraph conversion
up_down_matrix <- up_down_lsh %>%
  as.matrix()

# convert edge matrix to igraph
# note that directionality isn't attributed in the igraph object
# directionality is implicit when paths are created using upstream and downstream ends
# creating an igraph object without directionality allows for upstream path creation
test_igraph <- graph_from_edgelist(up_down_matrix, directed = FALSE)

# double check visually
#plot(test_igraph)
```

```{r}
# compute all simple paths between downstream ends and upstream ends
# and store them in a list of igraph paths
hold_paths <- lapply(downstream_ends,
                     function(x) all_simple_paths(test_igraph,
                                                  from = x,
                                                  to = upstream_ends))

# unplist igraph paths
hold_paths <- unlist(hold_paths, recursive = F)

# drop unused information like geodesic attributes from igraph paths
paths <- lapply(1:length(hold_paths), function(x) as_ids(hold_paths[[x]]))

# turn list of paths into dataframe
paths_df <- as.data.frame(as.matrix(paths)) %>%
  magrittr::set_colnames("paths_lst") %>%
  rowid_to_column(var = "row_id")

# create final dataframe of paths that has
# changing the `collapse` argument will change the separator in the path strings
paths_df_clean <- unnest(paths_df, paths_lst) %>%
  group_by(row_id) %>%
  mutate(paths_chr = paste0(paths_lst, collapse = " ")) %>%
  select(row_id, paths_chr) %>%
  unique() %>%
  mutate(downstream_end = str_sub(paths_chr, 1, 1),
         upstream_end = str_sub(paths_chr, -1, -1)) %>%
  left_join(paths_df, by = "row_id") %>%
  select(row_id, downstream_end, upstream_end, paths_chr, paths_lst)
```


```{r}
# take a look at the output dataframe
paths_df_clean
```
